---
title: "Stat632 HW6"
author: "Brandon Keck"
format: pdf
editor: visual
---

```{r,warning = FALSE, message=FALSE}
library(ggplot2)
library(readr)
county_votes16 <- read_csv("county_votes16.csv")
```

## Exercise 1.

### (a)

```{r}
glm1 <- glm(trump_win ~ obama_pctvotes, data = county_votes16, family = binomial)
summary(glm1)
```

Estimated Logistic Regression Equation:

$\boxed{\hat{p}(x) = \frac{e^{20.12971 - 0.37149x}}{1 + e^{20.12971 - 0.37149x}}}$

### (b)

```{r}
ggplot(county_votes16, aes(obama_pctvotes, trump_win)) + 
  geom_point() + 
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
  xlab("Percent Obama Votes") +
  ylab("Estimated Probability Trump Won") +
  theme_bw()
```

### (c)

```{r}
new_x <- data.frame(obama_pctvotes = c(40, 50, 60))

predict(glm1, newdata = new_x, type = "response")
```

40%: When Obama has 40% of the votes in a specific county, the estimated probability that Trump will win that county is 99.50%

50%: When Obama has 50% of the votes in a specific county, the estimated probability that Trump will win is 82.57%.

60%: When Obama has 60% of the votes in a specific county, the estimated probability that Trump will win that county is 10.35%.

### (d)

```{r}
exp(-0.37149)
```

For each 1 percentage point increase in Obama's 2012 vote share in a county, the log-odds of Trump winning the county in 2016 decreases by approximately 0.6897059.

## Exercise 2

### (a)

```{r}
glm2 <- glm(trump_win ~ pct_pop65 + pct_black + pct_white + pct_hispanic + pct_asian + highschool + bachelors + income, data = county_votes16, family = binomial)
summary(glm2)
```

### (b)

```{r}
glm3 <- glm(trump_win ~ pct_black + pct_white + pct_hispanic + pct_asian + highschool + bachelors + income, data = county_votes16, family = binomial)
summary(glm3)
```

### (c)

-   pct_black: (-0.0366) Counties with a higher percentage of Black residents are less likely to vote for Trump.

-   pct_white: (0.0819) Counties with a higher percentage of White residents are more likely to vote for Trump

-   pct_hispanic: (-0.0826) Higher Hispanic population is associated with lower likelihood of Trump winning

-   pct_asian: (-0.1521) Higher Asian population also predicts lower odds of Trump winning

-   highschool: (-0.0437) A higher percentage of residents with at least a high school diploma is slightly associated with decreased odds of Trump winning.

-   income: (0.0506) Higher per capita income is associated with a higher chance of Trump winning
