---
title: "HW 6"
author: "Brandon Keck"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 0: Leterature Review
https://www.ahajournals.org/doi/pdf/10.1161/HYPERTENSIONAHA.107.094011

#### 1. 
This article from the American Heart Association states that weight gain is associated with increases in systolic blood pressure, with each additional kilogram potentially raising blood pressure by 1-3 mm Hg. However, weight loss-even without dietary sodium adjustments can lead to measureable reductions in blood pressure.


#### 2.
Based upon the article that I read there appears to be an association between weight and blood pressure.As body weight increases, so too does systolic blood pressure.


## Part 1: Data Cleaning
#### 3.
```{r}
getwd()
```


```{r}
#load in the dataset. We are in the probs stats directory
chsdata <- read.table("chsData.txt", header = TRUE)
```



```{r}
# Convert clinic to a factor
chsdata$clinic <- factor(chsdata$clinic, 
                          levels = c(1, 2, 3, 4), 
                          labels = c("Sacramento", "Forsyth", "Washington", "Pittsburgh"))

```



```{r}
# Convert season to a factor
chsdata$season <- factor(chsdata$season,
                         levels = c(1, 2, 3, 4),
                         labels = c("summer", "fall", "winter", "spring"))
```


```{r}
# Convert arth to a factor
chsdata$arth <- factor(chsdata$arth,
                         levels = c(0, 1),
                         labels = c("none", "arthiritis"))
```



```{r}
chsdata$diab <- factor(chsdata$diab,
                         levels = c(1, 2, 3),
                         labels = c("none", "borderline", "diabetes"))
```



```{r}
# Convert income to a factor
chsdata$income <- factor(chsdata$income,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8),
                         labels = c("<=5k", "5k-8k", "8k-12k", "12k-16k", "16k-24k", "24k-35k", "35k-50k", ">50k"))
```


```{r}
chsdata$exint0 <- factor(chsdata$exint0, 
                          levels = c(0, 1, 2, 3), 
                          labels = c("no exercise", "low intensity", "moderate intensity", "high intensity"))
```



```{r}
head(chsdata)
```



#### (b)
```{r}
chsdata$sbp140 <- ifelse(chsdata$sbp >= 140, 1, 0)
```



## Part 2: Exploratory Data Analysis
#### (4a) 

```{r}
library(knitr)
```


```{r}
values <- sapply(chsdata, function(x) sum(is.na(x)))
```


```{r}
table <- data.frame(Variable = names(values), count = values)
```


```{r}
kable(table, colnames = c("Variable", "Number of Missing Rows"), caption = "Table of Missing Values for Each Variable")
```






