---
title: "Keck_Brandon_STAT620_HW6"
author: "Brandon Keck"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Section 4.1
##### Exercise 4.1-3

###### (a) Find fX (x), the marginal pmf of X. 

$$
f(x,y) = \frac{x+y}{32} , x = 1,2  \ and \ y = 1, 2, 3, 4.
$$


$$
f_x(x) = \sum_{y=1}^{4} 
$$


$$
= \frac{x+1}{32} + \frac{x+2}{32} + \frac{x+3}{32} + \frac{x+4}{32}
$$


$$
= \boxed{f_x(x)=\frac{4x+10}{32}, x = 1,2}
$$





###### (b)Find fY (y), the marginal pmf of Y. 


$$
f(x,y) = \frac{x+y}{32}, \ x = 1,2 \ and \ y=1,2,3,4.
$$


$$
f_y(y) = \sum_{x=1}^{2}
$$


$$
= \frac{1+y}{32} + \frac{2+y}{32}
$$


$$
= \boxed{f_y(y) = \frac{3+2y}{32}, \ y=1,2,3,4}
$$


###### (c) Find P(X > Y). 
The possible paris for x,y are (1,1), (1,2), (1,3), (1,4),(2,1), (2,2), (2,3) and (2,4)

The only pair that satisfies these conditions wher x > y is (2,1)

$$
P(X>Y) = \frac{2+1}{32}
$$

$$
=\boxed{\frac{3}{32}}
$$

###### (d) Find P(Y = 2X). 
The only pair that satisfies these conditions where Y = 2x is (1,2) and (2,4)


$$
\frac{1+2}{32} + \frac{2+4}{32}
$$

$$
\boxed{\frac{9}{32}}
$$



###### e Find P(X + Y = 3). 
The possible paris for x,y are (1,1), (1,2), (1,3), (1,4),(2,1), (2,2), (2,3) and (2,4)

(1,2), (2,1) satisfies this condition.

$$
\frac{1+2}{32} + \frac{2+1}{32}
$$


$$
\boxed{\frac{6}{32}}
$$


###### (f) Find P(X ≤ 3 − Y). 


The possible paris for x,y are (1,1), (1,2), (1,3), (1,4),(2,1), (2,2), (2,3) and (2,4)



$$
P(X \leq 3-Y)
$$


$$
P(X+Y \leq 3)
$$



The only pairs that satisfies these conditions is (1,1), (1,2) and (2,1)


$$
\frac{1+1}{32} + \frac{1+2}{32} + \frac{2+1}{32}
$$



$$
=\boxed{\frac{1}{4}}
$$




###### (g) Are X and Y independent or dependent? Why or why  not? 

Two discrete variables are said to be independent if P(X=x, Y=y) = P(X=x)P(Y=y) for all (x,y).

$$
f_x(x)= \frac{4x+10}{32} ,  \ x = 1,2
$$



$$
f_y(y)= \frac{3+2y}{32}, \ y = 1,2,3,4
$$



$$
f_x(x)*f_y(y) = \frac{4x+10}{32} * \frac{3+2y}{32}
$$




$$
= \frac{(4x+10) * (3+2y)}{1024}
$$



$$
f(x,y) \neq f_x(x) * f_y(y)
$$


Therefore, the random variables X and Y are dependent


###### (h) Find the means and the variances of X and Y. 

```{r}
x <- c(1, 2)
px <- c(14/32, 18/32)

sum(x * px)
```


```{r}
sum(x^2 * px) - (1.5625)^2
```



$$
\boxed{E(X)=1.5625, \ V(X) = 0.2460938}
$$


```{r}
y <- c(1, 2, 3, 4)
py <- c(5/32, 7/32, 9/32, 11/32)

sum(y * py)
```



```{r}
sum(y^2 * py) - (2.8125)^2
```



$$
\boxed{E(Y)= 2.8125 , \ V(Y) = 1.152344}
$$


\n

## Section 4.2
##### Exercise 4.2-9 f(x, y) = 1/4, (x, y) ∈ S = {(0, 0),(1, 1),(1, −1),(2, 0)}. 

###### (a) Are X and Y independent? 

To determine if X and Y are independet we need to check if each probability join event P(X=x, Y=y) = P(X=x)P(Y=y)

Marginal pmf of X

| X | Probability |
|:-:|:-----------:|
| 0 | $\frac{1}{4}$ |
| 1 | $\frac{1}{4}+ \frac{1}{4} = \frac{1}{2}$|
| 2 | $\frac{1}{4}$ |


\



$$
\frac{1}{4} + \frac{1}{2} + \frac{1}{4}
$$


$$
= 1 
$$


Therefore, this is a valid probability mass function because all values of probabilities are positive and add to 1


Marginal pmf of Y

| Y | Probability |
|:-:|:-----------:|
| -1 | $\frac{1}{4}$ |
| 0 | $\frac{1}{4} + \frac{1}{4}= \frac{1}{2}$ |
| 1 | $\frac{1}{4}$ |





$$
\frac{1}{4} + \frac{1}{2} + \frac{1}{4}
$$


$$
= 1 
$$


Therefore, this is a valid probability mass function because all values of probabilities are positive and add to 1.


$$
P(X=x, Y=y) = P(X=x)P(Y=y)
$$


$$
For \  (0,0) \ P(X=0, Y=0) = \frac{1}{4}
$$



$$
P(X=0)P(Y=0) = (\frac{1}{4}) (\frac{1}{2}) = \frac{1}{8}
$$


$$
Since \ \frac{1}{4} \neq \frac{1}{8}
$$

$$
\boxed{X \ and \ Y \ are \ dependent}
$$


###### (b) Calculate Cov(X, Y) and ρ. 

$$
\mu_x = E(X) = \sum x * P(X=x)
$$


$$
\mu_x = (0) * (\frac{1}{4}) + (1) * \frac{1}{2} + (2) * \frac{1}{4}
$$


$$
\mu_x= 1
$$




$$
\mu_y = E(Y) = \sum y * P(Y=y)
$$




$$
\mu_y = (-1) * (\frac{1}{4}) + (0) * \frac{1}{2} + (1) * \frac{1}{4}
$$




$$
\mu_y = 0
$$



$$
Cov(X,Y) = E(X,Y) - \mu_x \mu_y
$$



$$
(0)(0)(\frac{1}{4}) + (1)(1)(\frac{1}{4}) + (1)(-1)(\frac{1}{4}) + (2)(0)(\frac{1}{4}) - (1)(0)
$$



$$
\boxed{COV(X,Y) = 0}
$$



$$
\rho = \frac{Cov(X,Y)}{\sigma_X \sigma_Y} = \frac{\sigma_{XY}}{\sigma_X \sigma_Y}
$$



$$
E(X^2) = (0)^2 * (\frac{1}{4}) + (1)^2 * \frac{1}{2} + (2)^2 * \frac{1}{4}
$$


$$
E(X^2) = 1.5
$$


$$
\sigma^2_X = E(X^2)-[E(X)]^2
$$


$$
1.5 - (1^2)
$$


$$
= 0.5
$$


$$
E(Y^2) = (-1)^2 * (\frac{1}{4}) + (0)^2 * \frac{1}{2} + (1)^2 * \frac{1}{4}
$$


$$
E(Y^2) = \frac{1}{2}
$$



$$
\sigma^2_Y=E(Y^2)- [E(Y)]^2
$$




$$
0.5-(0^2)
$$



$$
=0.5
$$



$$
Corr(X,Y) = \frac{Cov(X,Y)}{\sqrt{var(X)}\sqrt{var(Y)}}
$$



$$
\frac{0}{\sqrt{0.5}\sqrt{0.5}}
$$



$$
\boxed{\rho = 0}
$$

\n

## Section 4.4
##### Exercise 4.4-2 Let X and Y have the joint pdf f(x, y) = x + y,  0 ≤ x ≤ 1, 0 ≤ y ≤ 1. 

###### (a)  

Marginal pdf of X:

$$
f_X(X) = \int_{-\infty}^{\infty} f(x,y)dy
$$



$$
\int_0^1 x+y \ dy
$$



$$
= xy +\frac{y^2}{2}]_0^1
$$


$$
x+ \frac{1}{2}
$$


$$
\boxed{f_X(X) = x + \frac{1}{2},\ 0 \le x \le 1}
$$


Marginal pdf of Y:
$$
f_Y(Y) = \int_{-\infty}^{\infty} f(X,Y)dx
$$



$$
\int_0^1 x + y \ dx
$$



$$
\frac{x^2}{2} + xy]_0^1
$$



$$
=y+ \frac{1}{2}
$$


$$
\boxed{f_Y(Y) = y + \frac{1}{2}, \ 0 \le y \le 1}
$$


$$
Two \ random \ variables \  x \ and \ y \ are \ independent \ if \ f_{X,Y}(X,Y)= f_{X}(X)f_{Y}(Y)
$$


$$
f_{X}(X)*f_{Y}(Y) = (x +\frac{1}{2}) * (y + \frac{1}{2})
$$



$$
= (xy + \frac{x}{2}+ \frac{y}{2}+\frac{1}{4})
$$



$$
x+y \neq (xy + \frac{x}{2}+ \frac{y}{2}+\frac{1}{4})
$$



$$
\boxed{X \ and \ Y \ dependent}
$$



###### (b)
###### (i)

$$
E(X) = \int_{-\infty}^{\infty} x * f_x(x) \ dx
$$



$$
E(X) = \int_{0}^{1} x * (x+\frac{1}{2}) \ dx
$$



$$
\int_{0}^{1} x^2+\frac{x}{2} dx
$$



$$
\frac{x^3}{3}+ \frac{x^2}{4}]_0^1
$$



$$
= \frac{1}{3}+ \frac{1}{4} = \frac{7}{12}
$$


$$
\boxed{\frac{7}{12}}
$$

###### (ii)

$$
E(Y)= \int_0^1 y * (y+ \frac{1}{2})dy
$$



$$
\int_0^1 y^2+\frac{y}{2}dy
$$


$$
\frac{y^3}{3} + \frac{y^2}{4}]_0^1
$$


$$
= \frac{1}{3}+ \frac{1}{4} = \frac{7}{12}
$$



$$
\boxed{\frac{7}{12}}
$$


###### (iii)
$$
E(X^2) = \int_0^1 x^2(x+\frac{1}{2})dx
$$

$$
\int_0^1 x^3 + \frac{x^2}{2}dx
$$


$$
\frac{x^4}{4}+ \frac{x^3}{6}]_0^1
$$


$$
\frac{1}{4}+ \frac{1}{6}= \frac{5}{12}
$$



$$
Var(X) = \frac{5}{12} - (\frac{7}{12})^2
$$


$$
\boxed{Var(X)=\frac{11}{144}}
$$


###### (iv)

$$
E(Y^2) = \int_0^1 y^2 * (y+\frac{1}{2})dy
$$


$$
\int_0^1 y^3 + \frac{y^2}{2}dy
$$


$$
\frac{y^4}{4}+ \frac{y^3}{6}]_0^1
$$


$$
\frac{1}{4}+ \frac{1}{6}= \frac{5}{12}
$$


$$
Var(Y) = \frac{5}{12} - (\frac{7}{12})^2
$$


$$
\boxed{Var(Y)= \frac{11}{144}}
$$

\n

##### Exercise 4.4-4
###### (a) Find P(0 ≤ X ≤ 1/2). 


Marginal pdf of X:

$$
f_X(X) = \int_{x^2}^1 \frac{3}{2}dy
$$


$$
\frac{3}{2}y]_{x^2}^1
$$


$$
f_X(X)=\frac{3}{2}- \frac{3}{2}x^2 , x^2 \le y \le 1
$$


Marginal pdf of Y:
$$
f_Y(Y)= \int_0^1 \frac{3}{2}dx
$$


$$
\frac{3}{2}x]_0^1
$$


$$
\frac{3}{2}(1)-\frac{3}{2}(0) = \frac{3}{2}
$$

$$
f_Y(Y) = \frac{3}{2} , 0 \le x \le 1
$$



$$
P(0 \le X \le \frac{1}{2})
$$


$$
\int_{0}^{\frac{1}{2}} \frac{3}{2}-\frac{3}{2}x^2dx
$$


$$
\frac{3}{2}x-\frac{3}{2} \frac{x^3}{3}]_0^{\frac{1}{2}}
$$


$$
= \frac{3}{2}(\frac{1}{2}) - \frac{1}{2}(\frac{1}{2})^3
$$


$$
=\boxed{\frac{11}{16}}
$$



###### (b) Find P(1/2 ≤ Y ≤ 1). 

$$
P(\frac{1}{2} \le Y \le 1) = \int_{\frac{1}{2}}^1 \frac{3}{2}dx
$$



$$
= \frac{3}{2}x]_{\frac{1}{2}}^1
$$


$$
\frac{3}{2}(1) - \frac{3}{2}(\frac{1}{2})
$$


$$
=\boxed {\frac{3}{4}}
$$


###### (c) Find P(X ≥ 1/2, Y ≥ 1/2). 

$$
P(X \ge \frac{1}{2}, Y \ge \frac{1}{2}) = \int_{\frac{1}{2}}^1 \int_{\frac{1}{2}}^1 \frac{3}{2}dxdy
$$


$$
= \int_{\frac{1}{2}}^1 \frac{3}{2}x]_{\frac{1}{2}}^1dy
$$


$$
\int_{\frac{1}{2}}^1 \frac{3}{2}(1)- \frac{3}{2}(\frac{1}{2})dy
$$


$$
= \int_{\frac{1}{2}}^1 \frac{3}{4}dy
$$


$$
\frac{3}{4}y]_{\frac{1}{2}}^1
$$



$$
= \frac{3}{4}(1) - \frac{3}{4}(\frac{1}{2})
$$


$$
=\boxed{\frac{3}{8}}
$$



###### (d) Are X and Y independent? Why or why not? 

$$
P(X=x, Y=y) = P(X=x)P(Y=y) \ for \ all \ x \ and \ y
$$


$$
f_X(X) * F_Y(Y) = (\frac{3}{2} - \frac{3}{2}x^2) * \frac{3}{2}
$$


$$
= \frac{9}{4}- \frac{9}{4}x^2
$$


$$
f(x,y) \neq f_X(X)* f_Y(Y)
$$


$$
Therefore, \ x \ and \ y \ are \ not \ independent.
$$

\n

##### Exercise 4.4-14
 
draw a graph that illustrates the domain of the pdf 
```{r}
# Define the x and y sequences
x <- seq(-0.2, 1.2, length = 50)
y <- seq(-0.2, 1.2, length = 50)
n <- length(x)

# Create an empty matrix for the function values
fxy <- matrix(0, nrow = n, ncol = n)

# Fill the matrix with the values of the function
for (i in 1:n) {
  for (j in 1:n) {
    if (x[i] >= 0 & x[i] <= 1 & y[j] >= 0 & y[j] <= 1 & x[i] <= y[j]) {  # Conditions for 0 <= x <= y <= 1
      fxy[i, j] <- 8 * x[i] * y[j]  # f(x, y) = 8xy
    }
  }
}

# Plot the 3D perspective plot
persp(x, y, fxy, zlim = c(0, 8), col = 'yellow', theta = -30, phi = 30, ticktype = "detailed",
      xlab = "X", ylab = "Y", zlab = "f(x, y)")

```

###### (a) Find the marginal pdfs of X and Y.

Marginal pdf of X, 
$$
f_X(X) = \int_{x}^1 8xy \ dy
$$


$$
= 8x \int_x^1 y \ dy
$$


$$
=8x [\frac{y^2}{2}]_x^1
$$


$$
= 8x(\frac{1}{2}- \frac{x^2}{2})
$$


$$
= 4x(1-x^2)
$$


Thus, the marginal pdf of X is:
$$
\boxed{f_X(X) = 4x(1-x^2), \ 0 \le x \le 1}
$$


The marginal pdf of Y,
$$
f_Y(Y) = \int_0^y 8xy \ dx
$$


$$
= 8y \int_0^yx \ dx
$$


$$
=8y [\frac{x^2}{2}]_0^y
$$


$$
= 8y * \frac{y^2}{2}
$$

$$
= 4y^3
$$


Thus, the marginal pdf of Y:
$$
\boxed{=f_Y(Y)=4y^3, \ 0 \le y \le 1}
$$


###### (b) Compute μX, μY, σ2 X, σ2 Y, Cov(X, Y), and ρ. 

$\mu_x:$


$$
\mu_x = E(X) = \int_0^1 x * f_X(X) \ dx
$$


$$
= \int_0^1 x * (4x(1-x^2) \ dx
$$


$$
= \int_0^1 x *(4x-4x^3) \ dx
$$


$$
= \int_0^1 4x^2 - 4x^4 \ dx
$$


$$
= \frac{4x^3}{3}- \frac{4x^5}{5}]_0^1
$$


$$
= \frac{4(1)^3}{3} - \frac{4(1)^5}{5}
$$


$$
= \frac{4}{3} - \frac{4}{5} = \frac{8}{15}
$$


$$
\boxed{\mu_X = \frac{8}{15}}
$$


$\mu_Y:$

$$
\mu_Y = E(Y) = \int_0^1 y * 4y^3 \ dy
$$


$$
=\int_0^1 4y^4 \ dy
$$


$$
= \frac{4y^5}{5}]_0^1
$$


$$
\boxed{\mu_Y =\frac{4}{5}}
$$


$\sigma^2_X = Var(X):$


$$
\sigma^2_X = E(X^2) - (\mu_X)^2
$$


$$
E(X^2) = \int_0^1 x^2 f_X(X) \ dx = \int_0^1 x^2 * (4x-4x^3) \ dx
$$


$$
= \int_0^1 4x^3-4x^5 \ dx
$$


$$
= \frac{4x^4}{4} - \frac{4x^6}{6}]_0^1
$$


$$
= \frac{4(1)^4}{4} - \frac{4(1)^6}{6} = \frac{4}{4} - \frac{4}{6}
$$


$$
= 1 - \frac{4}{6} = \frac{1}{3}
$$


$$
\sigma^2_X = E(X^2) - (\mu_X)^2 = \frac{1}{3} - (\frac{8}{15})^2
$$


$$
= \frac{1}{3} - \frac{64}{225} = \frac{11}{225}
$$


$$
Thus, \ \boxed{\sigma^2_X = \frac{11}{225}}
$$


$\sigma^2_Y = Var(Y):$

$$
\sigma^2_Y = E(Y^2)- (\mu_Y)^2
$$


$$
E(Y^2) = \int_0^1 y^2 * f_Y(Y) \ dy = \int_0^1 y^2 * 4y^3 \ dy
$$


$$
= \int_0^1 4y^5 \ dy
$$


$$
= \frac{4y^6}{6}]_0^1
$$


$$
= \frac{4}{6}
$$


$$
\sigma^2_Y = E(Y^2)-(\mu_Y)^2 = \frac{4}{6} - (\frac{4}{5})^2
$$


$$
=\frac{4}{6}-\frac{16}{25} = \frac{2}{75}
$$


$$
Thus, \ \boxed{\sigma^2_Y = \frac{2}{75}}
$$


$Cov(X,Y):$


$$
Cov(X,Y) = E(XY)- \mu_X\mu_Y
$$


$$
E(XY) = \int_0^1 xy * 8xy \ dxdy = \int_0^1 \int_0^y 8x^2y^2 \ dxdy
$$


$$
= \int_0^1 8y^2 \int_0^y x^2 dxdy = \int_0^1 8y^2 [\frac{x^3}{3}]_0^y dy
$$


$$
= \int_0^18y^2 * \frac{y^3}{3} dy = \int_0^1 \frac{8y^5}{3}dy
$$


$$
=\frac{8y^6}{18}]_0^1
$$


$$
\boxed{E(XY) = \frac{8}{18}}
$$


$$
Cov(X,Y) = E(XY) - \mu_X\mu_Y
$$


$$
= \frac{8}{18} - (\frac{8}{15})(\frac{4}{5}) 
$$


$$
\boxed{Cov(X,Y)= \frac{4}{225}}
$$


$\rho(Correlation \ Coefficient):$

$\rho = \frac{Cov(X,Y)}{\sigma_X \sigma_Y}$

```{r}
(4/225)/((sqrt(11/225))*(sqrt(2/75)))
```


$$
\boxed{\rho = 0.492366}
$$

###### (c)


$$
Y = \mu_Y + \frac{\rho \sigma_Y}{\sigma_X}(X - \mu_X)
$$


$$
\mu_X = \frac{8}{15}
$$


$$
\mu_Y = \frac{4}{5}
$$



$$
\sigma^2_X = \frac{11}{225} = \sigma_X = \sqrt{\frac{11}{225}}
$$


$$
\sigma^2_Y = \frac{2}{75} = \sigma_Y = \sqrt{\frac{2}{75}}
$$

$$
Cov(X,Y) = \frac{4}{225}
$$


$$
\sigma_X = \sqrt{\frac{11}{225}} = \frac{\sqrt{11}}{15}
$$


$$
\sigma_Y = \sqrt{\frac{2}{75}}= \frac{\sqrt2}{5\sqrt{3}} = \frac{\sqrt{6}}{15}
$$


$$
\rho = \frac{Cov(X,Y)}{\sigma_X\sigma_Y}
$$


$$
\rho = \frac{\frac{4}{225}}{\sqrt{\frac{11}{15}}*\frac{\sqrt{6}}{15}}
$$


$$
\frac{\frac{4}{225}}{\frac{\sqrt{66}}{225}} = \frac{4}{\sqrt{66}}
$$


$$
\frac{\rho \sigma_Y}{\sigma_X} = \frac{\frac{4}{\sqrt{66}}*\frac{\sqrt{6}}{15}}{\frac{\sqrt{11}}{15}}
$$


$$
\frac{4*\sqrt{6}}{\sqrt{66}* \sqrt{11}} = \frac{4}{11}
$$


$$
\boxed{Y = \frac{4}{5} + \frac{4}{11}(X - \frac{8}{15})}
$$






