---
title: "Keck_Brandon_STAT620_HW5"
author: "Brandon Keck"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 3.2-12 Let X equal the number of alpha particle emissions of carbon-14 that are counted by a Geiger counter  each second. Assume that the distribution of X is Poisson  with mean 16. Let W equal the time in seconds before the  seventh count is made. 

#### (a) Give the distribution of W. 

X is the number of alpha particle emissions counted each second and follows a Poisson distribution with mean 16. X ~ Poisson($\lambda$ = 16)

W follows a Gamma distribution with shape parameter $\alpha = 7$ and rate parameter $\lambda = 16$ 

$\boxed{W \sim Gamma(7,16)}$


#### (b) Find P(W ≤ 0.5). HINT: Use Equation 3.2-1 with  λw = 8. 

We are give that $\lambda w = 8$ and to use the Equation 3.2-1 which is

$$
F(w) = 1 - \sum_{k=0}^{\alpha-1} \frac{\lambda w^ke^{- \lambda w}}{k!}
$$


$$
1 - \sum_{k=0}^{7-1} \frac{8^ke^{-k}}{k!}
$$


```{r}
k <- c(0,1,2,3,4,5,6)

1 - sum(8^k * exp(-8) / factorial(k))
```

$\boxed{0.6866257}$

### 3.2-13. If X is χ2(23), find the following:  
#### (a) P(14.85 < X < 32.01). 

```{r}
pchisq(32.01, 23) - pchisq(14.85, 23)
```


$$
P(14.85 < X < 32.01) = \boxed{0.7999933}
$$

#### (b) Constants a and b such that P(a < X < b) = 0.95 and  P(X < a) = 0.025. 

```{r}
a <- qchisq(0.025, 23)
a

b <- qchisq(0.975, 23)
b
```

$$
a = \boxed{11.68855} 
$$

$$
b = \boxed {38.07563}
$$


#### (c) The mean and variance of X. 

The mean of the chi-square distribution is equal to the degrees of freedom which is r. The value of r is 23.

$$
\boxed{\mu = 23}
$$

The formula for the variance of a chi-square distribution is 
$$
\sigma^2 = 2r
$$

$$
2 * 23 = 46
$$

$$
\boxed{2r = 46}
$$





#### (d) χ2 0.05(23) and χ2 0.95(23). 

```{r}
qchisq(0.05, 23)
```


```{r}
qchisq(0.95, 23)
```


$$
\boxed{\chi^2_{0.05}= 13.09051}
$$


$$
\boxed{\chi^2_{0.95}=35.17246}
$$


### 3.3-5. If X is normally distributed with a mean of 6 and a variance of 25, find  

#### (a) P(6 ≤ X ≤ 14).

```{r}
pnorm(14, mean = 6, sd = 5) - pnorm(6, mean = 6, sd = 5)
```


$\boxed{0.4452007}$


#### (b) P(4 ≤ X ≤ 14). 

```{r}
pnorm(14, mean = 6, sd = 5) - pnorm(4, mean = 6, sd = 5)
```


$\boxed{0.6006224}$


#### (c) P(−4 < X ≤ 0).

```{r}
pnorm(0, mean = 6, sd = 5) - pnorm(-4, mean = 6, sd = 5)
```

$\boxed{0.09231954}$


#### (d) P(X > 15). 

```{r}
1 - pnorm(15, mean = 6, sd = 5)
```

$\boxed{0.03593032}$


#### (e) P(|X − 6| < 5) 
P(-5 < X -6 < 5) = P(1 < X < 11)

```{r}
pnorm(11, 6, sd = 5) - pnorm(1, 6, sd = 5)
```

$\boxed{0.6826895}$


#### (f) P(|X − 6| < 10). 

p(-10 < X-6 < 10) = P(-4 < X < 16)
```{r}
pnorm(16, 6, sd = 5) - pnorm(-4, 6, sd = 5)
```

$\boxed{0.9544997}$


#### (g) P(|X − 6| < 15).
P(-15 < X - 6 < 15) = P(-9 < X < 21)
```{r}
pnorm(21, mean = 6, sd = 5) - pnorm(-9, mean = 6, sd = 5)
```


 $\boxed{0.9973002}$
 
 
#### (h) P(|X − 6| < 12.4). 

```{r}
pnorm(18.4, 6, sqrt(25)) - pnorm(-6.4, 6, sqrt(25))
```

$\boxed{0.9868618}$

### 3.3-8. Let the distribution of X be N(μ, σ2). Show that  the points of inflection of the graph of the pdf of X occur  at x = μ ± σ. 


$$
f(x) = \frac{1}{\sigma \sqrt{2 \pi}}e^\frac{{-(x - \mu)^2}}{2 \sigma^2}
$$
$$
we \ must \ utilize \ the \ chain \ rule \ f'(g(x))*g'(x)
$$

$$
setting \ u = -\frac{(x- \mu)^2}{2 \sigma^2}
$$


$$
= -\frac{2(x- \mu)}{2 \sigma^2}
$$

$$
= \frac{x-\mu}{\sigma^2}
$$


$$
\frac{1}{\sigma \sqrt{2 \pi}}e^{-\frac{(x-\mu)^2}{2 \sigma^2}*u}
$$


$$
= \frac{1}{\sigma \sqrt{2 \pi}}e^{-\frac{(x-\mu)^2}{2 \sigma^2}* -\frac{x-\mu}{\sigma^2}}
$$

$$
= -\frac{x-\mu}{\sigma^3 \sqrt{2 \pi}}e^{-\frac{x-\mu}{2 \sigma^2}}
$$

Now we must take the derivative of f'(x) to get f''(x).

$$
u(x)= -\frac{(x- \mu)}{\sigma^3 \sqrt{2\pi}}
$$

$$
v(x) = e^{-\frac{(x-\mu)^2}{2\sigma^2}}
$$


$$
we \ can \ apply \ the \ product \ rule \ (fg)'=f'g+fg'
$$


$$
f''(x)=u'(x)*v(x)+u(x)*v'(x)
$$


$$
u'(x) = -\frac{1}{\sigma^3 \sqrt{2 \pi}}
$$

$$
v(x)= e^{-\frac{(x-\mu)^2}{2\sigma^2}}
$$


$$
u(x)= -\frac{x-\mu}{\sigma^3\sqrt{2\pi}}
$$


$$
v'(x)= e^{-\frac{(x-\mu)^2}{2\sigma^2}*{-\frac{x-\mu}{\sigma^2}}}
$$


$$
f''(x)= -\frac{1}{\sigma^3 \sqrt{2 \pi}} e^{-\frac{(x-\mu)^2}{2\sigma^2} + {-\frac{(x-\mu)^2}{2\sigma^2}}} * e^{-\frac{(x-\mu)^2}{2\sigma^2}*{-\frac{x-\mu}{\sigma^2}}}
$$


$$
f''(x) = \frac{e^{-\frac{(x-\mu)^2}{2\sigma^2}}}{\sigma^5\sqrt{2\pi}}(\sigma^2-(x-\mu)^2)
$$


$$
\frac{e^{-\frac{(x-\mu)^2}{2\sigma^2}}}{\sigma^5\sqrt{2\pi}}(\sigma^2-(x-\mu)^2) = 0
$$


$$
\sigma^2 - (x-\mu)^2=0
$$


$$
(x-\mu)^2= \pm \sigma^2
$$


$$
x-\mu = \pm \sigma
$$

$$
Therefore, \boxed{ \ x =\mu \pm\sigma}
$$


### 3.3-9. Find the distribution of W = X^2 when  
#### (a) X is N(0, 4), 

$$
\sigma = \sqrt{4}=2
$$

So X ~ N(0,4) X has a mean of 0 and a sd of 2


$$
f(x) = \frac{1}{2\sqrt{2\pi}}e^{-\frac{x^2}{8}}
$$


$$
F_W(w)=P(W \le w) = P(X^2 \le w)
$$


$$
P(X \le \pm \sqrt{w})
$$


$$
F_X(\pm \sqrt{w})
$$


$$
f_W(w)= \frac{dF_W(w)}{dw}
$$


$$
f_W(w)= \frac{1}{\sqrt{w}}*f_X(\sqrt{w})
$$


$$
f_X(x)= \frac{1}{2\sqrt{2\pi}}e^{-\frac{x^2}{8}}
$$


$$
\frac{1}{\sqrt{w}}* \frac{1}{2\sqrt{2\pi}}e^{-\frac{w}{8}}
$$


$$
f_W(w)= \frac{e^{-\frac{w}{8}}w^{\frac{1}{2}-1}}{\Gamma{\frac{1}{2}}*8^{\frac{1}{2}}}
$$


$$
Thus, \ W \ follws \ a \ gamma \ distribution \ with \ parameters \ \alpha = \frac{1}{2} \ and \ \theta = 8
$$


#### (b) X is N(0, σ2). 

$$
F_W(w)=P(W \le w) = P(X^2 \le w)
$$

$$
P(X \le \pm \sqrt{w})
$$


$$
F_X(\pm \sqrt{w})
$$


$$
f_W(w)= \frac{dF_W(w)}{dw}
$$


$$
f_W(w)= \frac{1}{\sqrt{w}}*f_X(\sqrt{w})
$$


$$
f_X(x)= \frac{1}{2\sqrt{2\pi}}e^{-\frac{x^2}{8}}
$$

$$
\frac{1}{\sqrt{w}}* \frac{1}{2\sqrt{2\pi}}e^{-\frac{w}{8}}
$$


$$
f_W(w)= \frac{e^{-\frac{w}{8}}w^{\frac{1}{2}-1}}{\Gamma{\frac{1}{2}}*\sqrt{2\sigma^2}}
$$


$\boxed{W \sim\Gamma({\frac{1}{2},2\sigma^2})}$




### 3.3-11. A candy maker produces mints that have a label  weight of 20.4 grams. Assume that the distribution of the  weights of these mints is N(21.37, 0.16). 

#### (a) Let X denote the weight of a single mint selected at  random from the production line. Find P(X > 22.07). 

```{r}
1 - pnorm(22.07, mean = 21.37, sd = 0.4)
```

$\boxed{0.04005916}$


#### (b) Suppose that 15 mints are selected independently and  weighed. Let Y equal the number of these mints that  weigh less than 20.857 grams. Find P(Y ≤ 2). 


Find the probability that a single mints weighs less than 20.857 grams
```{r}
pnorm(20.857, 21.37, 0.4)
```




```{r}
pbinom(2, size = 15, prob = 0.09983365)
```
 

$\boxed{0.8165799}$