---
title: "HW 8"
author: "Brandon Keck"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 5.1

### 5.1-6

#### (a) Find the cdf of X
$$
f(x) = 2x, \ 0 < x < 1
$$

$$
F(x) = \int_0^x 2t dt
$$


$$
=t^2]_0^x
$$


$$
=x^2
$$

CDF F(x) is:
F(x) = 
\begin{cases} 
0 & \text{if } x \le 0, \\
x^2 & \text{if } 0 < x < 1, \\
1 & \text{if } x \ge 1.
\end{cases}


#### (b)
We can generate a uniform random variable of U on the interval (0,1)
Then we can compute that $X = \sqrt{U}$ which will give a sample from the desired distribution.


#### (c)
```{r}
set.seed(11101987) # Scorpio!!!

U <- runif(10)
X_sim <- sqrt(U)

X_sim
```

### 5.1-10
The lifetime (in years) of a manufactured product  is Y = 5X^0.7, where X has an exponential distribution  with mean 1. Find the cdf and pdf of Y. 


Since X is exponential with mean 1, its PDF is:
$$
f_X(x) = e^{-x}, x \ge 0
$$


CDF is:
$$
F_X(x)= 1 - e^{-x}, x \ge 0
$$

Derive the CDF of Y
$$
F_Y(y) = P(Y \le y)
$$


$$
= P(5X^{0.7} \le y)
$$


$$
= P(X^{0.7} \le \frac{y}{5})
$$


$$
=P(X \le (\frac{y}{5})^{\frac{1}{0.7}})
$$

Express in terms of $F_X(x)$
$$
F_Y(y) = F_X((\frac{y}{5})^{\frac{1}{0.7}})
$$


CDF of X:
$$
F_Y(y)= 1 - e^{-(\frac{1}{0.7})}, y \ge 0
$$


Derive PDF of Y:
$$
f_Y(y) = \frac{d}{dy}(1-e^{-(\frac{y}{5})^{\frac{1}{0.7}}})
$$


$$
f_Y(y)=e^{-(\frac{y}{5})^{\frac{1}{0.7}}} * \frac{d}{dy} (-(\frac{y}{5})^{\frac{1}{0.7}})
$$


$$
=e^{-(\frac{y}{5})^{\frac{1}{0.7}}} * \frac{1}{5*0.7} * (\frac{y}{5})^{\frac{1}{0.7}-1}
$$


$$
=\boxed{\frac{1}{3.5} * (\frac{y}{5})^{\frac{-0.3}{0.7}} * e^{-(\frac{y}{5})^{\frac{1}{0.7}}}, y \ge 0}
$$


### 5.1-17
#### (a)
The PDF of a standard normal distribution is:
$$
f(x) = \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}
$$


$$
X^2 = Y
$$



$$
X = - \sqrt{Y}, \ -1 < x < 1
$$

Derivative with respect to Y
$$
\frac{d}{dY} - \sqrt{Y}
$$


$$
=\boxed{-\frac{1}{2\sqrt{Y}}}
$$



$$
X = \sqrt{Y}, \ 0 < x < 1
$$


Derivative with respect to Y:
$$
\frac{d}{dY} \sqrt{Y}
$$



$$
=\boxed{ \frac{1}{2\sqrt{Y}}}
$$


#### (b)
$$
F_Y(y) = \int_{-\sqrt{y}}^{\sqrt{y}} \frac{3}{2}x^2 \ dx
$$


$$
= \frac{x^3}{3}]_{-\sqrt{y}}^{\sqrt{y}}
$$


$$
3[\frac{x^3}{3}]_0^{\sqrt{y}}
$$


$$
=\sqrt{y}^3 = y^{\frac{3}{2}}
$$


The CDF of Y:
$$
F_Y(y) = y^{\frac{3}{2}}, \ 0 \le y \le 1
$$

$$
f_Y(y)= \frac{d}{dy}(y^\frac{3}{2}) = \frac{3}{2}y^{\frac{1}{2}}
$$


$$
f_Y(y) = \boxed{\frac{3}{2}y^{\frac{1}{2}} , \ 0 \le y \le 1}
$$


## Section 5.3
### 5.3-2
#### (a)

```{r}
dbinom(2, size = 3, prob = 0.5)

dbinom(4, size = 5, prob = 0.5)
```



$$
P(Y_1 = 2, Y_2 = 4)
$$



```{r}
0.375 * 0.15625
```


$$
\boxed{0.05859375}
$$


#### (b)
$$
P(X_1+X_2 = 7)
$$


$$
X_1 + X_2 \sim b(3+5, 1/2) = b(8,1/2)
$$


```{r}
dbinom(7, size = 8, prob = 0.5)
```


$$
\boxed{0.03125}
$$



### 5.3-3
#### (a)
$$
P(0.5 < X_1 < 1) = \int_{0.5}^1 2x_1 dx_1
$$


$$
= x_1^2
$$

$$
P(0.5 < X_1 < 1) = [x_1^2]_{0.5}^1
$$


$$
= 1^2-(0.5)^2 = 1 - 0.25 = 0.75
$$


$$
P(0.5 < X_1 < 1) = 0.75
$$


$$
P(0.4 < X_2 < 0.8) = \int_{0.4}^{0.8} 4x_2^3 \ dx_2
$$


$$
= x_2^4
$$


$$
P(0.4 < X_2 < 0.8) = [x_2^4]_{0.4}^{0.8}
$$


$$
= (0.8)^4 - (0.4)^4
$$


```{r}
(0.8)^4 - (0.4)^4
```




```{r}
0.75 * 0.384
```



$$
P(0.5 < X_1 < 1 \ and \ 0.4 < X_2 < 0.8)
$$


$$
\boxed{0.288}
$$


#### (b)

$$
E(X_1^2) = \int_0^1 x_1^2 * f_1(x_1) dx_1
$$


$$
= \int_0^1 x_1^2 * 2x_1 dx_1
$$



$$
= \int_0^1 2x_1^3 dx_1
$$


$$
= \frac{2x_1^4}{4} = \frac{x_1^4}{2}
$$


$$
E(X_1) = [\frac{x_1^4}{2}]_{0}^{1} = \frac{1}{2}- 0
$$


$$
E(X_2^3) = \int_0^1 x_2^3 *f_2(x_2)dx_2
$$


$$
= \int_0^1 x_2^3 * 4x_2^3 dx_2
$$


$$
=\int_0^1 4x_2^6 dx_2
$$


$$
= \frac{4x_2^7}{7} = \frac{4}{7}x_2^7
$$


$$
= \frac{4}{7}-0 = \frac{4}{7}
$$


```{r}
0.5 * 4/7
```


$$
\boxed{E(X_1^2X_2^3) = 0.2857143}
$$



### 5.3-14
The exponential distribution
$$
f(y) = \frac{1}{\lambda}e^{-\frac{1}{\lambda}y}
$$



$$
F(y) = P(Y \le k)
$$


$$
=1-P(Y > k)
$$


$$
\int_0^y \frac{1}{\lambda}e^{-\frac{k}{\lambda}}dk
$$


$$
= [\frac{1}{\lambda} \frac{e^{-\frac{k}{\lambda}}}{-\frac{1}{\lambda}}]_0^y
$$



$$
=[-e^{-\frac{k}{\lambda}}]_0^y
$$


$$
=(1-e^{-\frac{y}{\lambda}})
$$


$$
=(e^{-\frac{y}{\lambda_1}})(e^{-\frac{y}{\lambda_2}})(e^{-\frac{y}{\lambda_3}})
$$





$$
=(e^{-\frac{1000}{1000}})(e^{-\frac{1000}{1500}})(e^{-\frac{1000}{2000}})
$$


```{r}
# Calculate each probability term
exp(-1000 / 1000) * exp(-1000 / 1500) * exp(-1000 / 2000)

```



$$
\boxed{0.1145588}
$$


## Section 5.4
### 5.4-2

$$
X \sim b(n,p)
$$


$$
M_X(t) = (1-p + pe^t)^n
$$


$$
Since \ X_1 \sim b(n_1,p)
$$


$$
M_X1(t) = (1-p + pe^t)^{n1}
$$


$$
M_X2(t) = (1-p + pe^t)^{n2}
$$


$$
M_Y(t) = M_X1(t) * M_X2(t)
$$


$$
M_Y(t) = (1-p + pe^t)^{n1} * (1-p + pe^t)^{n2}
$$


$$
M_Y(t) = (1-p + pe^t)^{n1+n2}
$$



The mgf $M_Y(t) = (1-p+pe^t)^{n1+n2}$ is a binomial random variable with parameters $n = n_1 + n_2$

$$
\boxed{M_Y(t)= (1-p + pe^t)^{n1+n2}}
$$

$$
\boxed{Y \sim b(n_1+n_2)}
$$


### 5.4-20

$$
P(Y > 90) = \int_{90}^{\infty} f_Y(x)dx
$$



$$
f_Y(x) = \frac{x^{6-1}e^{-\frac{x}{10}}}{\Gamma(6)*10^6}
$$



$$
P(Y > 90) = \int_{90}^{\infty} \frac{x^5e^{-\frac{x}{10}}}{\Gamma(6)*10^6} dx
$$



```{r}
1 - pgamma(90, shape = 6, scale = 10)
```

$$
\boxed{0.1156905}
$$



## Section 5.5
### 5.5-2
#### (a)
```{r}
curve(dnorm(x, mean = 50, sd = 6), from = 20, to = 80)
```



#### (b)
$Var(\bar{X}) = \frac{36}{9} = 4$


$SD(\bar{X}) = \sqrt{4}=2$


$Thus, \bar{X} \ follows \ a \ normal \ distribution: \bar{X} \sim N(50,4)$


```{r}
curve(dnorm(x, mean = 50, sd = 2), from = 40, to = 60)
```



#### (c)

$\bar{X} \sim N(50,1)$


```{r}
curve(dnorm(x, mean = 50, sd = 1), from = 45, to = 55)
```


### 5.5-5

$\mu = 8.78 \ and \ \sigma = 0.4$


$\mu \pm 3\sigma = 8.78 \pm3 * 0.4$

$= (7.58, 9.98)$

$Variance \ of \ \bar{X} = \frac{0.16}{9}$

$\sigma = \sqrt{0.01777}$

$=0.1333$


$\mu  \pm 3\sigma = 8.78 \pm 3 * 0.1333$

$=(8.38,9.2)$

```{r}
# Plot the PDF of X 
curve(dnorm(x, mean = 8.78, sd = 0.4), from = 7.5, to = 10,
      col = adjustcolor("skyblue", alpha.f = 0.5), lwd = 1.5,
      xlab = "x", ylab = "Density",
      main = "PDFs of X and Sample Mean",
      ylim = c(0, 3))  # Set y-axis limits from 0 to 3

# Add the PDF 
curve(dnorm(x, mean = 8.78, sd = 0.1333), from = 7.5, to = 10,
      col = adjustcolor("red", alpha.f = 0.5), lwd = 1.5,
      lty = 2, add = TRUE)

# Add a legend to differentiate 
legend("topright", legend = c("X ~ N(8.78, 0.16)", "X̄ ~ N(8.78, 0.0167)"),
       col = c(adjustcolor("skyblue", alpha.f = 0.5), adjustcolor("red", alpha.f = 0.5)), 
       lwd = 1.5, lty = c(1, 2))
```




