---
title: "HW 9"
author: "Brandon Keck"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 5.6

## 5.6-1
The formula for the mean of the uniform distribution is
$$
\mu = \frac{a+b}{2}
$$


$$
\mu = \frac{0+1}{2} = \frac{1}{2}
$$
The variance formula is
$$
\sigma^2 = \frac{(b-a)^2}{12}
$$

$$
\sigma^2 = \frac{(1-0)^2}{12} = \frac{1}{12}
$$

$$
\frac{\frac{1}{\sqrt{12}}}{\sqrt{12}} = \frac{1}{12}
$$

z-score formula
$$
z = \frac{\bar{x}-\mu_{\bar{x}}}{\sigma_x}
$$


$$
z_1 = \frac{\frac{1}{2}-\frac{1}{2}}{\frac{1}{12}} = 0
$$

$$
z_2 = \frac{\frac{2}{3}-\frac{1}{2}}{\frac{1}{12}} = 2
$$

$$
P(\frac{1}{2} \le \bar{x} \le \frac{2}{3}) = P(z_1 < 2) - P(z_2 <0)
$$

```{r}
0.9772 - 0.5000
```

$$
\boxed{0.4772}
$$


## 5.6-8
### a.
The central limit theorem states that the sampling distribution of the sample mean is approximately normal with mean $\mu_x$ and standard deviation $\sigma_x$

$\mu_\bar{x} = \mu$

Thus $\boxed{\mu_\bar{x} = 24.43}$

### b.
The formula for variance of $\bar{x} \ is \ \sigma_\bar{x}^2 = \frac{\sigma^2}{n}$

```{r}
2.20/30
```


$\boxed{\sigma_\bar{x}^2 = 0.07333333}$



### c.
z-score formula
$$
z = \frac{\bar{x}-\mu_{\bar{x}}}{\sigma_x}
$$


$z_1$
```{r}
(24.17 - 24.43) / (sqrt(0.07333333))
```

$z_2$
```{r}
(24.82 - 24.43) / (sqrt(0.07333333))
```


$$
P(24.17 \le \bar{x} \le 24.82)
$$


```{r}
0.9251 - 0.1685
```


$\boxed{0.7566}$

## 5.6-14
Formula for mean of X is:
$$
\mu_X = n * \mu
$$

```{r}
20 * 10
```


Formula for standard deviation is 
$$
\sigma_X = \sqrt{n} * \sigma
$$

```{r}
sqrt(20) * 2
```


```{r}
1- 0.2
```


The value of z-score from the table is 0.84
In order to get X
X = $\mu + z * \sigma$

```{r}
200 + 0.84 * 8.944272
```

$\boxed{207.5132}$

# Section 5.7
## 5.7-4
$$
P(35 \le X \le 40) = P(34.5 < X < 40.5)
$$

$$
=P(\frac{34.5 - 36}{\sqrt{9}} < \frac{X-36}{\sqrt{9}} < \frac{40.5-36}{\sqrt{9}})
$$


$$
=P(-0.5 < W < 1.50)
$$

$$
=P(W < 1.50) - P(W < -0.50)
$$

```{r}
0.9332 - 0.3085
```


$\boxed{0.6247}$


## 5.7-9
### a.

$$
P(15 < \sum_{i=1}^{30} X \le 22) = P(15 < \sum_{i=1}^{30} X < 22.5)
$$

$$
=P(\frac{15-30 * \frac{2}{3}}{\sqrt{30 * \frac{2}{3}}}) < \frac{\sum_{i=1}^{30}X-30 * \frac{2}{3}}{\sqrt{30 * \frac{2}{3}}} < \frac{22.5-30 * \frac{2}{3}}{\sqrt{30 * \frac{2}{3}}}
$$

$$
=P(\frac{-5}{\sqrt{20}} < Y < \frac{2.5}{\sqrt{20}})
$$

$$
=P(Y < 0.56) - P(Y < -1.01)
$$

```{r}
0.7123-0.1562
```

$\boxed{0.5561}$

### b.

$$
=P(\frac{20.5-30 * \frac{2}{3}}{\sqrt{30 * \frac{2}{3}}}) < \frac{\sum_{i=1}^{30}X-30 * \frac{2}{3}}{\sqrt{30 * \frac{2}{3}}} < \frac{26.5-30 * \frac{2}{3}}{\sqrt{30 * \frac{2}{3}}}
$$

$$
=P(0.11 < Y < 1.45)
$$

$\boxed{0.3823}$

## 5.7-16 
p=0.1
$$
P(-1.5 < Y - 10 < 1.5) \approx \phi(\frac{1.5}{3}) - \phi(-\frac{1.5}{3})
$$

```{r}
0.6915 - 0.3085
```

$\boxed{0.383}$


p=0.5
$$
P(-1.5 < Y - 10 < 1.5) \approx \phi(\frac{1.5}{5}) - \phi(-\frac{1.5}{5})
$$
```{r}
0.6179 - 0.3821
```

$\boxed{0.2358}$


p=0.8
$$
P(-1.5 < Y - 10 < 1.5) \approx \phi(\frac{1.5}{4}) - \phi(-\frac{1.5}{4})
$$

```{r}
0.6462 - 0.3538
```

$\boxed{0.2924}$


# Section 5.8
## 5.8-1
### a.
Chebyshev's inequality
$$
P(|X - \mu| \ge k\sigma) \le \frac{1}{k^2}
$$

Given:
$$
\mu =33
$$

$$
\sigma^2 = 16
$$

$$
P(23 < X < 43) = P(|X-33| < 10)
$$

$$
k = \frac{10}{\sigma} = \frac{10}{4} = 2.5
$$

$$
P(|X - \mu| < 10) \ge 1 -  \frac{1}{(2.5)^2}
$$

$$
P(|X - \mu| < 10) \ge 1 -  \frac{1}{6.25}
$$

$\boxed{0.84}$

### b.
$$
k = \frac{14}{\sigma} = \frac{14}{4} = 3.5
$$

$$
P(|X - 33| \ge 14) \le \frac{1}{(3.5)^2}
$$

$$
P(|X - 33| \ge 14) \le \frac{1}{(12.25)}
$$

$$
P(|X - 33| \ge 14) \le 0.0816
$$

$\boxed{0.0816}$

## 5.8-6
$$
P(75 < \bar{X} < 85)
$$

$$
E(\bar{X}) = \mu = 80
$$

$$
Var(\bar{X}) = \frac{\sigma^2}{n} = \frac{60}{15} = 4
$$


$$
k = \frac{5}{\sigma_{\bar{X}}} = \frac{5}{2} = 2.5
$$

$$
P(|\bar{X} - \mu| < k\sigma_{\bar{X}}) \ge 1 -  \frac{1}{(k)^2}
$$

$$
P(|\bar{X} - 80| < 5) \ge 1 -  \frac{1}{(2.5)^2}
$$

$$
P(|\bar{X} - 80| < 5) \ge 1 -  \frac{1}{6.25}
$$


$\boxed{0.84}$


# Section 5.9
## 5.9-3

$$
M_{S^2}(t) = M_{\frac{\sigma^2}{n-1}\chi^2_{n-1}}(t)
$$

$$
= M_{\chi^2_{n-1}}(\frac{\sigma^2}{n-1}*t)
$$

$$
= (1-2 * \frac{\sigma^2}{n-1}*t)^{-(\frac{n-1}{2})}
$$

$$
= (1 + \frac{\frac{\sigma^2}{-(n-1)}}{2} * t) ^ {-(\frac{n-1}{2})}
$$

$$
lim_{n->\infty} M_{S^2}(t) = lim_{n->\infty} (1 + \frac{\frac{\sigma^2}{-(n-1)}}{2}*t) ^ {-\frac{(n-1)}{2}}
$$


$$
= (1-\frac{2t\sigma^2}{n-1}) ^{\frac{-(n-1)}{2}}
$$

$$
= e^{\sigma^2t}
$$


$$
Therefore, \ \boxed{lim_{n->\infty} (1+\frac{x}{n})^n = e^x}
$$

## 5.9-4


$$
Y = X_1 + X_2 + ....+X_n
$$

$$
X_i \approx \chi^2_1
$$

$$
E(X_i)=1
$$


$$
V(X_i) = 2
$$



$$
E(Y) = E(\sum_{i=1}^{n}X_i)
$$

$$
=\sum_{i=1}^{n}E(X_i)
$$

$$
= \sum_{i=1}^{n}1 = n
$$

$$
V(Y) = V(\sum_{i=1}^{n}X_i)
$$

$$
= \sum_{i=1}^{n}V(X_i)
$$


$$
= \sum_{i=1}^{n}2 = 2n
$$


$$
W = \frac{Y-E(Y)}{\sqrt{V(Y)}}
$$

$$
= \frac{Y-n}{\sqrt{2n}}
$$

$$
=\boxed{\frac{Y-n}{\sqrt{2n}} \sim N(0,1)}
$$

